---
title: <center style="font-size:30px;font-style:normal;color:black;">Supplementary Code 3:</center>
subtitle: <center style="font-size:30px;font-style:normal;color:#0E0E7D;">Mantel Tests & Spatial Autocorrelation</center>
 &nbsp;
author: |
  <center style="font-style:normal;">
  <a style="font-size:22px;color:#337ab7;text-decoration: underline;"href="http://smith-vidaurre.com/"> Grace Smith-Vidaurre</a><sup><span style="font-size:12px;color:black;text-decoration:none!important;">1*</span></sup>, 
  <a style="font-size:22px;color:#337ab7;text-decoration: underline;"href="http://marceloarayasalas.weebly.com">Marcelo Araya-Salas</a><sup><span style="font-size:12px;color:black;text-decoration:none!important;">2</span></sup>, 
  <a style="font-size:22px;color:#337ab7;text-decoration: underline;"href="http://wrightbehaviorlab.org">Timothy F. Wright</a><sup><span style="font-size:12px;color:black;text-decoration:none!important;">1</span></sup></center>
  &nbsp;
  <center style="font-size:18px;font-style:normal;color:black;"><sup><span style="font-size:12px;color:black;">1</span></sup>Department of Biology, New Mexico State University, <sup><span style="font-size:12px;color:black;">2</span></sup>Recinto de Golfito, Universidad de Costa Rica</center>
  <br />
  <center style="font-size:18px;"><sup style="font-size:12px;">*</sup>gsvidaurre@gmail.com</center>
  &nbsp;
date: <center style="font-size:22px;font-style:normal;>`r format(Sys.time(), '%d %B, %Y')`</center>
  <br />
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

<style type="text/css">

a:hover {
  color: #23527c !important;
}

h1.title {
  font-size: 32px;
  color: black;
  font-weight: normal;
}

h1 {
   color: black;
   font-size: 26px;
   font-weight: normal;
}

h2 {
   color: black;
   font-size: 24px;
   font-weight: bold;
}

h3 {
   color: black;
   font-size: 20px;
   font-weight: normal;
}

h4 {
   color: black;
   font-size: 20px;
   font-weight: normal;
}

body{ /* Normal */
      font-size: 18px;
  }
code.r{ /* Code block */
    font-size: 18px;
}
</style>

We evaluated patterns of acoustic similarity over social scales and geographic distance. We used Mantel tests<a href='#References'><sup>[1]</sup></a> to ask whether contact calls were more similar within social groups compared to among social groups at each social scale. We used Mantel tests and Mantel-based spatial autocorrelation to evaluate whether acoustic similarity decreased linearly over increasing geographic distance. 

Our overall results can be reproduced by using the data provided with this publication (extended selection tables for the individual and higher social scales, and selection table spreadsheets). Some changes to this code may be necessary, such as directories used for analysis.

We ran sound analysis in Supplementary Methods 2 to calculate acoustic similarity by SPCC and random forests. As before, we sometimes use the naming convention "site-level" to refer to calls for higher social scales. 
```{r setup, eval = TRUE, echo = FALSE}

knitr::opts_knit$set(root.dir = "/media/owner/MYIOPSITTA/R/Uruguay2017_MonkParakeet_CallAnalysis/")

```

```{r echo = TRUE, eval = TRUE, message = FALSE}

rm(list = ls())

X <- c("ggplot2", "vegan", "pbapply", "parallel", "data.table", "facetscales", "rgeos", "rgdal", "sp", "tidyverse")

invisible(lapply(X, library, character.only = TRUE))

```

```{r echo = TRUE, eval = TRUE}

path <- "/media/owner/MYIOPSITTA/R/Uruguay2017_MonkParakeet_CallAnalysis/Data"

```

Pre-processed repeatedly sampled individual selection table and metadata.
```{r echo = TRUE, eval = TRUE}

ccs_fi <- read.csv(file.path(path, "Focal_Individuals/indiv_scale_calls_preprocessed.csv"), header = TRUE)

```

Pre-processed higher social scale call selection table and metadata.
```{r echo = TRUE, eval = TRUE}

ccs <- read.csv(file.path(path, "Site_Recordings/higher_social_scales_calls_preprocessed_final.csv"), header = TRUE)

ccs$call_ID <- paste(ccs$sound.files, ccs$selec, sep = "-")

```

# Pairwise Geographic Distance

Project geographic coordinates and calculate pairwise distance among sites at different geographic scales.
```{r echo = TRUE, eval = FALSE}

# need to convert to Spatial Points object
mat <- as.matrix(data.frame(lon = ccs$lon, lat = ccs$lat))
sp_pts <- SpatialPoints(mat, proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# access EPSG codes to reproject in meters
# used EPSG 5383 for Uruguay
epsg <- rgdal::make_EPSG()
# str(epsg)

epsg[grep("^5383$", epsg$code), ]

# reproject
sp_pts <- sp::spTransform(sp_pts, CRSobj = CRS(epsg$prj4[grep("^5383$", epsg$code)]))
# bbox(sp_pts)
# proj4string(sp_pts)

# calculate pairwise distances among all sites
geo_dists <- rgeos::gDistance(sp_pts, byid = TRUE)
# str(geo_dists)

# repeat for Colonia
Colonia <- ccs[grep("Colonia", ccs$dept), ]

mat <- as.matrix(data.frame(lon = Colonia$lon, lat = Colonia$lat))
sp_pts <- SpatialPoints(mat, proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# reproject
sp_pts <- sp::spTransform(sp_pts, CRSobj = CRS(epsg$prj4[grep("^5383$", epsg$code)]))
# bbox(sp_pts)
# proj4string(sp_pts)

# calculate pairwise distances among all sites
geo_distsC <- rgeos::gDistance(sp_pts, byid = TRUE)
# str(geo_distsC)

# repeat for all sites except Salto
tmp <- droplevels(data.frame(lat = ccs$lat[-grep("ARAP", ccs$General_site)], lon = ccs$lon[-grep("ARAP", ccs$General_site)], site = ccs$General_site[-grep("ARAP", ccs$General_site)]))

mat <- as.matrix(data.frame(lon = tmp$lon, lat = tmp$lat))
sp_pts <- SpatialPoints(mat, proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# reproject
sp_pts <- sp::spTransform(sp_pts, CRSobj = CRS(epsg$prj4[grep("^5383$", epsg$code)]))
# bbox(sp_pts)
# proj4string(sp_pts)

# calculate pairwise distances among all sites
geo_dists2 <- rgeos::gDistance(sp_pts, byid = TRUE)
# str(geo_dists2)

path <- "/media/owner/MYIOPSITTA/R/Uruguay2017_MonkParakeet_CallAnalysis/Data/Site_Recordings"

saveRDS(geo_dists, file.path(path, "geo_dists.RDS"))
saveRDS(geo_dists2, file.path(path, "geo_dists2.RDS"))
saveRDS(geo_distsC, file.path(path, "geo_distsC.RDS"))

```

```{r echo = TRUE, eval = TRUE}

path <- "/media/owner/MYIOPSITTA/R/Uruguay2017_MonkParakeet_CallAnalysis/Data/Site_Recordings"

geo_dists <- readRDS(file.path(path, "geo_dists.RDS"))
geo_dists2 <- readRDS(file.path(path, "geo_dists2.RDS"))
geo_distsC <- readRDS(file.path(path, "geo_distsC.RDS"))

```

## Summary Statistics

We generated summary statistics of geographic distances in the full and Colonia site data sets.

Mean and range of pairwise geographic distances among all sites? In kilometers.
```{r echo = TRUE, eval = TRUE}

mean(geo_dists)/1000

range(geo_dists[geo_dists != 0])/1000 

# all sites except ARAP in the Salto department
mean(geo_dists2)/1000

range(geo_dists2[geo_dists2 != 0])/1000 

```

Mean and range of pairwise geographic distances among Colonia sites? Also in kilometers.
```{r echo = TRUE, eval = TRUE}

mean(geo_distsC)/1000

range(geo_distsC[geo_distsC != 0])/1000 

```

# Similarity Matrices

We prepared similarity matrices for analysis.
```{r echo = TRUE, eval = TRUE}

path <- "/media/owner/MYIOPSITTA/R/Uruguay2017_MonkParakeet_CallAnalysis/Data"

xc_mat <- readRDS(file.path(path, "Site_Recordings/xc_mat_higher_social_scales.RDS"))

proxm_site <- readRDS(file.path(path, "Site_Recordings/proxm_site.RDS"))

xc_mat_fi <- readRDS(file.path(path, "Focal_Individuals/xc_mat_indiv_scale.RDS"))

```

```{r echo = TRUE, eval = TRUE}

# site calls for site, flock and pair social scales
xc_dist <- stats::as.dist(1 - xc_mat, diag = TRUE, upper = TRUE)
# str(xcdist)

rf_dist <- stats::as.dist(1 - proxm_site, diag = TRUE, upper = TRUE)
# str(rf_dist)

# repeatedly sampled individual calls for individual scale
xcdist_fi <- stats::as.dist(1 - xc_mat_fi, diag = TRUE, upper = TRUE)
# str(xcdist_fi)

```

We prepared additional data for Mantel tests, including a call ID variable.
```{r echo = TRUE, eval = TRUE}

# site calls
ccs$call_ID <- paste(ccs$sound.files, ccs$selec, sep = "-")

ccs$call_ID <- sapply(1:nrow(ccs), function(x){
  tmp <- strsplit(ccs$call_ID[x], split = "_")[[1]]
  paste(tmp[4], tmp[5], sep = "_")
})

# change call_ID for better grepping later
ccs$call_ID <- gsub(".WAV", "", ccs$call_ID)
ccs$call_ID <- gsub("-", "_", ccs$call_ID)

# repeated individual calls
ccs_fi$call_ID <- paste(ccs_fi$sound.files, ccs_fi$selec, sep = "-")

ccs_fi$site <- as.character(ccs_fi$site)

```

# General Approach

Mantel tests encompassed 4 social scales, 2 geographic scales (regional: all sites across the transect and local: sites in the Colonia department) and 2 acoustic similarity methods. We encoded pairwise social group membership at each social scale by generating pairwise binary identity matrices (e.g. 1 = the same social group, 0 = different social groups). We converted acoustic similarity and binary identity matrices to distances matrices for Mantel tests by subtracting matrices from 1. 

# Mantel Tests

```{r echo = TRUE, eval = TRUE}

path <- "/media/owner/MYIOPSITTA/R/Uruguay2017_MonkParakeet_CallAnalysis/Data/Site_Recordings"

geo_dists <- readRDS(file.path(path, "geo_dists.RDS"))
geo_distsC <- readRDS(file.path(path, "geo_distsC.RDS"))

```

```{r echo = TRUE, eval = FALSE}

# cores for parallel processing
cores <- parallel::detectCores() - 2

# permutations for Mantel tests
perms <- 9999

# loop over call data sets
call_data_set <- c("higher social scale", "Individual")

# loop over acoustic similarity methods for the site call data set, these are the response matrices for Mantel tests
acoustic_sim <- c("SPCC", "RF")

# will convert these to distance matrices inside the loop, these matrices have a dimension name structure that is easier to parse
response <- list(xc_mat, proxm_site)

# loop over social scales for the site call data set
# will create binary identity matrices per social scale within the loop to serve as predictors
type <- c("Site", "Site - Colonia", "Flock", "Pair")

# convert geographic distance matrices to km
# these matrices will also serve as predictors
geo_dists <- geo_dists/1000
geo_distsC <- geo_distsC/1000

# read in pair and flock IDs from Supplementary Methods 1
pair_IDs <- readRDS(file.path(path, "pair_IDs.RDS"))
length(pair_IDs)

flock_IDs <- readRDS(file.path(path, "flock_IDs.RDS"))
length(flock_IDs)

# create an empty csv file that will be populated by appending Mantel test results
file.remove(file.path(path, "Mantel_results.txt")) # remove previous versions

invisible(pblapply(1:length(call_data_set), function(d){
  
  # split up Mantel tests by call data set
  # higher social scale calls will be used for multiple tests at different social and geographic scales
  if(call_data_set[d] == "Site-level"){
    
    # loop over acoustic similarity methods
    lapply(1:length(acoustic_sim), function(a){
      
      # loop over test type (social or geographic scales) 
      lapply(1:length(type), function(y){
        
          # cat(paste(d, a, y, "\n", sep = ""))
        
        # subset acoustic similarity response matrices and predictor matrices as needed, not necessary for site social and geographic scale
        if(grepl("^Site$", type[y])){
          
          tmp <- ccs
          
          ####################################################
          # Prep acoustic response matrix
          resp <- response[[a]]
          resp_tmp <- resp

          # convert to a distance matrix
          resp_tmp_dist <- stats::as.dist(1 - resp_tmp, diag = TRUE, upper = TRUE)
          # dim(resp_tmp_dist)
          
          ####################################################
          # Prep predictor matrix of site membership for Mantel tests
          id_sites <- unlist(lapply(1:nrow(tmp), function(i){
              return(as.numeric(tmp$General_site[i] == tmp$General_site))
          }))
           # str(id_sites)

          id_sites_mat <- matrix(id_sites, nrow = nrow(tmp), ncol = nrow(tmp), dimnames = list(tmp$call_ID, tmp$call_ID))

          id_sites_dist <- stats::as.dist(1 - id_sites_mat, diag = TRUE, upper = TRUE)
          # str(id_sites_dist)
            
          ####################################################
          # Run Mantel tests for site membership and geographic distance
          mntl1 <- vegan::mantel(xdis = id_sites_dist, ydis = resp_tmp_dist, method = "pearson", permutations = perms, parallel = cores)
            # str(mntl1)
             
          # create csv file for first Mantel iteration
          if(acoustic_sim[a] == "SPCC"){
            write.table(data.frame(social_scale = type[y], acoustic_similarity_method = acoustic_sim[a], geo_dist = "N", partial_Mantel = "N", num_calls = dim(id_sites_mat)[1], mantel_r = round(mntl1$statistic, digits = 2), mantel_p = mntl1$signif, permutations = perms), file = file.path(path, "Mantel_results.txt"), sep = ",", col.names = TRUE, quote = FALSE, row.names = FALSE)
            
          # if now on RF method, append results rather than writing a new csv
          } else if(acoustic_sim[a] == "RF"){
          write.table(data.frame(social_scale = type[y], acoustic_similarity_method = acoustic_sim[a], geo_dist = "N", partial_Mantel = "N", num_calls = dim(id_sites_mat)[1], mantel_r = round(mntl1$statistic, digits = 2), mantel_p = mntl1$signif, permutations = perms), file = file.path(path, "Mantel_results.txt"), sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE, append = TRUE)
        }
            
          # give the geographic distance matrix unique dimension names
          # dimnames(geo_dists) <- tmp$call_ID
          attributes(geo_dists)$Labels <- tmp$call_ID
          mntl2 <- vegan::mantel(xdis = geo_dists, ydis = resp_tmp_dist, method = "pearson", permutations = perms, parallel = cores)
          # str(mntl2)
              
           write.table(data.frame(social_scale = type[y], acoustic_similarity_method = acoustic_sim[a], geo_dist = "Y", partial_Mantel = "N", num_calls = dim(id_sites_mat)[1], mantel_r = round(mntl2$statistic, digits = 2), mantel_p = mntl2$signif, permutations = perms), file = file.path(path, "Mantel_results.txt"), append = TRUE, sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE)
  
        }
        
        else if(!grepl("^Site$", type[y])){
            
         if(grepl("Colonia|Flock|Pair", type[y])){
            
            # initialize column for indexing call data set
            if(type[y] == "Site - Colonia") col_nm <- "dept"
            if(type[y] == "Flock") col_nm <- "Flock_ID"
            if(type[y] == "Pair") col_nm <- "Pair_ID"
            
            # indices of calls for the given social scale in the higher social scale data set
            if(grepl("Colonia", type[y])){
              ss <- which(ccs[[col_nm]] == "Colonia")
            } else if(grepl("Flock|Pair", type[y])){
              
              if(type[y] == "Pair"){
                gids <- pair_IDs
              } else if(type[y] == "Flock"){
                gids <- flock_IDs
              }
    
              ss <- grep(paste(paste("^", gids, "$", sep = ""), collapse = "|"), ccs[[col_nm]])

            }
             
            # calls for pairs or flocks in the higher social scale data set
            ss_calls <- ccs$call_ID[ss] 
            
            # subset call metadata 
            tmp <- ccs[grep(paste(paste("^", ss_calls, "$", sep = ""), collapse = "|"), ccs$call_ID), ]
            # unique(tmp$call_ID == ss_calls) # checking
            
            ####################################################
            # Filter acoustic similarity response matrix for the given social scale and acoustic similarity method
  # initialize response matrix depending on the current acoustic similarity method 
            resp <- response[[a]]
            length(dimnames(resp))

            # reassign row and column names for parsing
            dimnames(resp) <- list(ccs$call_ID, ccs$call_ID)

            pat <- paste(paste("^", ss_calls, "$", sep = ""), collapse = "|")
            # filter the response matrix by the calls of interest 
            resp_tmp <- resp[grep(pat, dimnames(resp)[[1]]), grep(pat, dimnames(resp)[[2]])]
            # str(resp_tmp)
            
            # convert to a distance matrix
            resp_tmp_dist <- stats::as.dist(1 - resp_tmp, diag = TRUE, upper = TRUE)
            # checking the filtering was performed correctly
            # unique(dimnames(resp_tmp_dist) == ss_calls)
            # unique(dimnames(resp_tmp_dist) == ss_calls)
            # unique(dimnames(resp_tmp_dist) == tmp$call_ID)
            # unique(dimnames(resp_tmp_dist) == tmp$call_ID)
            
            # cat(paste("Finished", type[y], "resp_tmp_dist", "\n", sep = " "))
            
            ####################################################
            # Generate binary identity matrices of social group membership
            
            if(grepl("Colonia", type[y])){
            
              id <- unlist(lapply(1:nrow(tmp), function(i){
                  return(as.numeric(tmp[["General_site"]][i] == tmp[["General_site"]]))
              }))
              
              # give the geographic distance matrix unique dimension names
              # dimnames(geo_distsC) <- tmp$call_ID
              attributes(geo_distsC)$Labels <- tmp$call_ID
              
            } else if(grepl("Flock|Pair", type[y])){
              
              id <- unlist(lapply(1:nrow(tmp), function(i){
                  return(id <- as.numeric(tmp[[col_nm]][i] == tmp[[col_nm]]))
              }))
            }
            
            # turn id vector into a matrix with the correct dimensions
            id_mat <- matrix(id, nrow = nrow(tmp), ncol = nrow(tmp), dimnames = list(tmp$call_ID, tmp$call_ID))
            # str(id_mat)

            # convert this matrix to a distance matrix
            id_dist <- stats::as.dist(1 - id_mat, diag = TRUE, upper = TRUE)
            # str(id_dist)

            ####################################################
            # Run full Mantel tests for sites in Colonia
            # Site membership and geographic distance
            # Write out results as separate csv files
            if(grepl("Colonia", type[y])){
              mntl1 <- vegan::mantel(xdis = id_dist, ydis = resp_tmp_dist, method = "pearson", permutations = perms, parallel = cores)
              
              write.table(data.frame(social_scale = type[y], acoustic_similarity_method = acoustic_sim[a], geo_dist = "N", partial_Mantel = "N", num_calls = dim(id_mat)[1], mantel_r = round(mntl1$statistic, digits = 2), mantel_p = mntl1$signif, permutations = perms), file = file.path(path, "Mantel_results.txt"), append = TRUE, sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE)
              
              # first give the geo. dist. matrix unique dimnames
              # dimnames(geo_distsC) <- tmp$call_ID
              attributes(geo_distsC)$Labels <- tmp$call_ID
              mntl2 <- vegan::mantel(xdis = geo_distsC, ydis = resp_tmp_dist, method = "pearson", permutations = perms, parallel = cores)
              
              write.table(data.frame(social_scale = type[y], acoustic_similarity_method = acoustic_sim[a], geo_dist = "Y", partial_Mantel = "N", num_calls = dim(id_mat)[1], mantel_r = round(mntl2$statistic, digits = 2), mantel_p = mntl2$signif, permutations = perms), file = file.path(path, "Mantel_results.txt"), append = TRUE, sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE)
              
              
            } else if(grepl("Flock|Pair", type[y])){
            
            # make a site identity matrix for flock and pair calls 
            # will serve as the third matrix in the partial Mantel
            id_sites <- unlist(lapply(1:nrow(tmp), function(i){
              return(as.numeric(tmp$General_site[i] == tmp$General_site))
            }))
            # str(id_sites)

            id_sites_mat <- matrix(id_sites, nrow = nrow(tmp), ncol = nrow(tmp), dimnames = list(tmp$call_ID, tmp$call_ID))

            id_sites_dist <- stats::as.dist(1 - id_sites_mat, diag = TRUE, upper = TRUE)
            # str(id_sites_dist)
            
            ####################################################
            # Run partial Mantel tests for flocks and pairs
            # These scales are nested within sites
            mntl <- vegan::mantel.partial(xdis = id_dist, ydis = resp_tmp_dist, zdis = id_sites_dist, method = "pearson", permutations = perms, parallel = cores)
            
            write.table(data.frame(social_scale = type[y], acoustic_similarity_method = acoustic_sim[a], geo_dist = "N", partial_Mantel = "Y", num_calls = dim(id_mat)[1], mantel_r = round(mntl$statistic, digits = 2), mantel_p = mntl$signif, permutations = perms), file = file.path(path, "Mantel_results.txt"), append = TRUE, sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE)
              
            }
            
          } 
        } 
    })
  })
} else if(call_data_set[d] == "Individual"){
    
    ####################################################
    # acoustic response matrix (SPCC) generated previously, as were coordinates for site-pair table
    tmp <- ccs_fi
    resp_tmp <- xc_mat_fi
    # str(resp_tmp)
    
    # convert to a distance matrix
    resp_tmp_dist <- stats::as.dist(1 - resp_tmp, diag = TRUE, upper = TRUE)
    
    ####################################################
    # Make matrices of individual and site identity (individuals are nested within sites), prep for Mantel tests
    id <- unlist(lapply(1:nrow(tmp), function(i){
      return(as.numeric(tmp$indiv[i] == tmp$indiv))
    }))

    id_mat <- matrix(id, nrow = nrow(tmp), ncol = nrow(tmp), dimnames = list(tmp$call_ID, tmp$call_ID))
    # id_mat

    id_dist <- stats::as.dist(1 - id_mat, diag = TRUE, upper = TRUE)
    
    id_sites <- unlist(lapply(1:nrow(tmp), function(i){
      return(as.numeric(tmp$site[i] == tmp$site))
    }))
    # str(id_sites)

    id_sites_mat <- matrix(id_sites, nrow = nrow(tmp), ncol = nrow(tmp), dimnames = list(tmp$call_ID, tmp$call_ID))

    id_sites_dist <- stats::as.dist(1 - id_sites_mat, diag = TRUE, upper = TRUE)
    # str(id_sites_dist)
    
    ####################################################
    # Run Mantel test
    mntl <- vegan::mantel.partial(xdis = id_dist, ydis = resp_tmp_dist, zdis = id_sites_dist, method = "pearson", permutations = perms, parallel = cores)
    # str(mntl)
              
    write.table(data.frame(social_scale = call_data_set[d], acoustic_similarity_method = "SPCC", geo_dist = "N", partial_Mantel = "Y", num_calls = dim(id_mat)[1], mantel_r = round(mntl$statistic, digits = 2), mantel_p = mntl$signif, permutations = perms), file = file.path(path, "Mantel_results.txt"), append = TRUE, sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE)

  }
  
}))

```

Read in Mantel results and perform Bonferroni p-value correction for multiple comparisons. 
```{r echo = TRUE, eval = FALSE}

mantel_res <- read.table(file.path(path, "Mantel_results.txt"), sep = ",", header = TRUE)
str(mantel_res)

# write this out as a .csv
write.csv(mantel_res, file.path(path, "Mantel_results.csv"), row.names = FALSE)

```

```{r echo = TRUE, eval = TRUE}

mantel_res <- read.csv(file.path(path, "Mantel_results.csv"), header = TRUE)

# the total number of comparisons for Mantel tests:
nrow(mantel_res)

# the adjusted alpha by a Bonferroni correction:
alpha <- 0.05

adj_alpha <- round(alpha/nrow(mantel_res), digits = 4)
adj_alpha

```

See Table II in the main body of the article for Mantel test results.

# Spatial Autocorrelation

We used Mantel-based spatial autocorrelation to further evaluate the relationship between contact call similarity and geographic distance. In an individual recongition system, individuals at geographically distant sites might converge on similar calls, as their chances of social interaction would be quite low. Spatial autocorrelation could therefore provide insight not only about the nature of the relationship between call similarity and geographic distance, but also the minimum distance at which calls became more similar or different than expected. 
```{r echo = TRUE, eval = TRUE}

geo_dists <- readRDS(file.path(path, "geo_dists.RDS"))
geo_distsC <- readRDS(file.path(path, "geo_distsC.RDS"))

```

## Regional Scale

```{r echo = TRUE, eval = TRUE}

# recalculate geographic distance matrix among all site calls
tmp <- data.frame(lat = ccs$lat, lon = ccs$lon, site = ccs$General_site)
geo_dists <- round(geo_dists/1000) # convert to km and round values to integers

# find meaningful breaks in the geographic distance matrix
geo_vect <- geo_dists[lower.tri(geo_dists)]
geo_vect <- geo_vect[geo_vect != 1 & !is.na(geo_vect)]
range(geo_vect)

# various classes have 0 observations with a break interval of 5km
geo_classes <- cut(geo_vect, breaks = seq(0, round(max(geo_vect)), 5))
table(geo_classes)

# only 2 classes have 0 observations with a break interval of 10km
geo_classes <- cut(geo_vect, breaks = seq(0, round(max(geo_vect)), 10))
table(geo_classes)
dim(table(geo_classes))

# no classes have 0 observations with a break interval of 25km
geo_classes <- cut(geo_vect, breaks = seq(0, round(max(geo_vect)), 25))
table(geo_classes)

# choose a break value and define geographic distance classes for correlogram
classes <- seq(0, max(geo_dists), 10)
length(classes)

```

```{r echo = TRUE, eval = FALSE}

# Mantel's correlation using Pearson's method
correl_SPCC_10km <- vegan::mantel.correlog(D.eco = xc_dist, D.geo = geo_dists, break.pts = classes, cutoff = FALSE, r.type = "pearson", nperm = 999, mult = "holm", progressive = TRUE)
str(correl_SPCC_10km)

saveRDS(correl_SPCC_10km, file.path(path, "correl_SPCC_10km.RDS"))

correl_RF_10km <- vegan::mantel.correlog(D.eco = rf_dist, D.geo = geo_dists, break.pts = classes, cutoff = FALSE, r.type = "pearson", nperm = 999, mult = "holm", progressive = TRUE)
str(correl_RF_10km)

saveRDS(correl_RF_10km, file.path(path, "correl_RF_10km.RDS"))

```

## Local Scale

We repeated this analysis over a local geographic scale (Colonia department)
```{r echo = TRUE, eval = TRUE}

# Colonia sites
Colonia <- ccs[grep("Colonia", ccs$dept), ]
tmp <- data.frame(lat = Colonia$lat, lon = Colonia$lon, site = Colonia$General_site)

# convert to km and round values to integers
geo_distsC <- round(geo_distsC/1000) 

# find meaningful breaks in the geographic distance matrix
geo_vect <- geo_distsC[lower.tri(geo_distsC)]
geo_vect <- geo_vect[geo_vect != 1 & !is.na(geo_vect)]
range(geo_vect)

# 9 classes have 0 observations with a break interval of 2km
geo_classes <- cut(geo_vect, breaks = seq(0, round(max(geo_vect)), 2))
table(geo_classes)

# only 2 classes have 0 observations with a break interval of 5km
geo_classes <- cut(geo_vect, breaks = seq(0, round(max(geo_vect)), 5))
table(geo_classes)

# subset SPCC matrix by Colonia calls
xc_mat_Col <- xc_mat
dimnames(xc_mat_Col) <- list(as.character(ccs$dept), as.character(ccs$dept))
indices <- grep("^Colonia$", dimnames(xc_mat_Col)[[1]])
xc_mat_Col <- xc_mat_Col[indices, indices]
dim(xc_mat_Col)

# repeat for RF matrix
proxm_site_Col <- proxm_site
dimnames(proxm_site_Col) <- list(as.character(ccs$dept), as.character(ccs$dept))
indices <- grep("^Colonia$", dimnames(proxm_site_Col)[[1]])
proxm_site_Col <- proxm_site_Col[indices, indices]
dim(proxm_site_Col)

# convert both to distance matrices
xc_dist_Col <- stats::as.dist(1 - xc_mat_Col, upper = TRUE, diag = TRUE)
rf_dist_Col <- stats::as.dist(1 - proxm_site_Col, upper = TRUE, diag = TRUE)

# proceeding with classes at 2km intervals, since native range colonies are quite often within 2km 
classes <- seq(0, max(geo_distsC), 2)
length(classes)

```

```{r echo = TRUE, eval = FALSE}

# Mantel's correlation using Pearson's method
correl_SPCC_Col_2km <- vegan::mantel.correlog(D.eco = xc_dist_Col, D.geo = geo_distsC, break.pts = classes, cutoff = FALSE, r.type = "pearson", nperm = 999, mult = "holm", progressive = TRUE)
str(correl_SPCC_Col_2km)
saveRDS(correl_SPCC_Col_2km, file.path(path, "correl_SPCC_Col_2km.RDS"))

correl_RF_Col_2km <- vegan::mantel.correlog(D.eco = rf_dist_Col, D.geo = geo_distsC, break.pts = classes, cutoff = FALSE, r.type = "pearson", nperm = 999, mult = "holm", progressive = TRUE)
str(correl_RF_Col_2km)
saveRDS(correl_RF_Col_2km, file.path(path, "correl_RF_Col_2km.RDS"))

```

```{r echo = TRUE, eval = TRUE}

path <- "/media/owner/MYIOPSITTA/R/Uruguay2017_MonkParakeet_CallAnalysis/Data/Site_Recordings/"

geo_dists <- readRDS(file.path(path, "geo_dists.RDS"))
geo_distsC <- readRDS(file.path(path, "geo_distsC.RDS"))

geo_dists <- geo_dists/1000
geo_distsC <- geo_distsC/1000

correl_SPCC_Col_2km <- readRDS(file.path(path, "correl_SPCC_Col_2km.RDS"))
correl_RF_Col_2km <- readRDS(file.path(path, "correl_RF_Col_2km.RDS"))

correl_SPCC_10km <- readRDS(file.path(path, "correl_SPCC_10km.RDS"))
correl_RF_10km <- readRDS(file.path(path, "correl_RF_10km.RDS"))

```

#### Figure 4: Mantel-based Correlogram

This is Figure 4 in the main manuscript. 
```{r echo = TRUE, eval = TRUE, dpi = 200, fig.width = 7, fig.height = 5}

# make a data frame per acoustic similarity method and geographic scale
df1 <- as.data.frame(correl_SPCC_10km$mantel.res)
df2 <- as.data.frame(correl_RF_10km$mantel.res)
df3 <- as.data.frame(correl_SPCC_Col_2km$mantel.res)
df4 <- as.data.frame(correl_RF_Col_2km$mantel.res)

# remove NAs in corrected p-values (distance classes that had 0 or too few observations)
df1 <- df1[!is.na(df1$`Pr(corrected)`), ]
df2 <- df2[!is.na(df2$`Pr(corrected)`), ]
df3 <- df3[!is.na(df3$`Pr(corrected)`), ]
df4 <- df4[!is.na(df4$`Pr(corrected)`), ]

# combine distance classes
distance <- c(df1$class.index, df2$class.index, df3$class.index, df4$class.index)

# combine Mantel correlation values
corr <- c(df1$Mantel.cor, df2$Mantel.cor, df3$Mantel.cor, df4$Mantel.cor)

# combine corrected p-values
p_val <- c(df1$`Pr(corrected)`, df2$`Pr(corrected)`, df3$`Pr(corrected)`, df4$`Pr(corrected)`)

# make a vector of significance for plotting
sig <- c(ifelse(df1$`Pr(corrected)` <= 0.05, "sig", "non-sig"), ifelse(df2$`Pr(corrected)` <= 0.05, "sig", "non-sig"), ifelse(df3$`Pr(corrected)` <= 0.05, "sig", "non-sig"), ifelse(df4$`Pr(corrected)` <= 0.05, "sig", "non-sig"))

# use the vector of statistical significance together with Mantel correlation values to make a new vector of acoustic similarity as: 1) less than expected, 2) neutral and 3) higher than expected
sig_cat <- sig
sig_cat[sig_cat == "sig" & corr < 0] <- "Lower"
sig_cat[sig_cat == "sig" & corr > 0] <- "Higher"
sig_cat[sig_cat == "non-sig"] <- "Neutral"
unique(sig_cat)

# acoustic similarity method
type <- c(rep("SPCC", nrow(df1)), rep("Random_Forests", nrow(df2)), rep("SPCC", nrow(df3)), rep("Random_Forests", nrow(df4)))

# geographic scale
geo_scale <- c(rep("Regional", nrow(df1)), rep("Regional", nrow(df2)), rep("Local", nrow(df3)), rep("Local", nrow(df4)))

# combine variables for ggplotting
corr_df <- data.frame(distance = distance, corr = corr, p_val = p_val, sig = sig, sig_cat = sig_cat, type = type, geo_scale = geo_scale)

# organize factor levels as desired for plotting
corr_df$type <- factor(corr_df$type, levels = c("SPCC", "Random_Forests"))
corr_df$geo_scale <- factor(corr_df$geo_scale, levels = c("Local", "Regional"))
corr_df$sig <- factor(corr_df$sig, levels = c("sig", "non-sig"))
corr_df$sig_cat <- factor(corr_df$sig_cat, levels = c("Lower", "Neutral", "Higher"))

# intitialize x-axis scales for facetscales::facet_grid_sc
scales_x <- list(
    Local = scale_x_continuous(limits = c(0, 22), breaks = seq(0, 22, 2), labels = seq(0, 22, 2)),
    
  Regional = scale_x_continuous(limits = c(0, 470), breaks = seq(0, 470, 40), labels = seq(0, 470, 40))
)

# intitialize y-axis scales for facetscales::facet_grid_sc
scales_y <- list(
  SPCC = scale_y_continuous(limits = c(-0.16, 0.16)),
  
  Random_Forests = scale_y_continuous(limits = c(-0.16, 0.16))
)

ggplot(corr_df, aes(x = distance, y = corr)) +
  
  geom_hline(yintercept = 0, linetype = "dotted", size = 0.5) +
  
  geom_point(aes(color = sig_cat, fill = sig_cat), size = 1.75, shape = 21) +
  
  # scale_color_manual(values = c(alpha("darkblue", 0.6), gray.colors(12)[6], "red"), labels = c("Lower", expression("p "*">"*" 0.05"), "Higher")) +

  scale_color_manual(values = c(alpha("darkblue", 0.6), gray.colors(12)[6], "red")) +
  
  scale_fill_manual(values = c(alpha("darkblue", 0.6), gray.colors(12)[6], "red")) +
  
  geom_line(colour = "black", size = 0.25) +
  
  facetscales::facet_grid_sc(cols = vars(geo_scale), rows = vars(type), scales = list(x = scales_x, y = scales_y)) +

  theme_bw() +
  
  theme(panel.background = element_rect(color = "white", fill = "white"), panel.grid.major.x = element_line(size = 0.1, color = "black"), axis.line = element_line(color = "black", size = 0.35), axis.text.x = element_text(size = 6, angle = 0, face = "bold"), axis.title = element_text(size = 10), axis.ticks = element_line(size = 0.15), legend.text = element_text(hjust = 0, size = 10), legend.title = element_text(hjust = 0, size = 10), panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(), legend.position = "top", strip.text = element_text(size = 10)) +
  
  xlab("Pairwise Geographic Distance (km)") + ylab("Mantel Spatial Correlation of Acoustic Similarity") + 
  
  guides(fill = guide_legend(title = "Spatial Correlation Values"), color = guide_legend(title = "Spatial Correlation Values"))

# ggsave(file.path(path, "Figure4_Mantel_correlogram.tiff"), width = 5.00, height = 4.5, units = "in", dpi = 325)

```

Which distance classes were significant for SPCC at the regional and local scales?
```{r echo = TRUE, eval = TRUE}

# subtract 5 from the minimum distance and add 5 to the maximum distance to get the full interval of distances (binned by 10km)
# here these bins encompass 60 - 80km
corr_df %>% 
  filter(sig_cat == "Lower" & type == "SPCC" & geo_scale == "Regional")

corr_df %>% 
  filter(sig_cat == "Higher" & type == "SPCC" & geo_scale == "Local")
corr_df %>% 
  filter(sig_cat == "Lower" & type == "SPCC" & geo_scale == "Local")

```

Which distance classes were significant for RF at the regional and local scales?
```{r echo = TRUE, eval = TRUE}

# the first set of bins encompasses 60 - 110km
corr_df %>% 
  filter(sig_cat == "Lower" & type == "Random_Forests" & geo_scale == "Regional")

corr_df %>% 
  filter(sig_cat == "Higher" & type == "Random_Forests" & geo_scale == "Local")

corr_df %>% 
  filter(sig_cat == "Lower" & type == "Random_Forests" & geo_scale == "Local")

```

# References

    1. Mantel, N. (1967). The detection of disease clustering and a generalized regression approach. Cancer Research, 27(2), 209–220.
